### 2021-12-06 input/proposal.tex
rework start of CP

This definition encompasses more specific definitions of related
work~\cite{hsiao2005scoped, lue2006scoped, wang2015prodiluvian}.  It
also highlights that the functioning of epidemic propagation is
well-aligned with the objectives of scoped broadcast. More precisely,
it underlines the transitive relevance of messages, thus a \process
can stop \TODO{forwarding messages} as soon as its predicate becomes
unverified.  We use \NAMEB to efficiently modify the state of each
\process depending on the partitions that exist in the system.

@@

At any time, a \process can decide to become a \emph{source}, hence
creating a new partition in the system by executing an \texttt{Add}
operation. This partition includes at least its source plus
neighboring \processes that estimate they are closer to this source
than any other one. Such a distance (or \emph{weight}) is
application-dependant: in the context of maintaining distributed
indexes, this would be about link latency that \nodes could monitor by
aggregating \texttt{ping}s; or operational costs when dealing with
multiple tenants.

@@

\begin{definition}[\label{def:partitioning}Partitioning]
  Let $S \subseteq V$ be the set of \underline{s}ources, and $P_{s\in
    S}$ be the \underline{p}artition including at least \Process~$s$,
  each \process belongs to at most one partition $\forall x, y \in V,
  \forall s,s' \in S: (x \in P_{s} \wedge y \in P_{s'}) \implies (x \neq
  y \vee s = s')$, and there exists at least one path $\pi_{xs}$ of
  \processes that belong to this partition $\forall z \in \pi_{xs}: z
  \in P_s$.
\end{definition}

Definition~\ref{def:scoped} and Definition~\ref{def:partitioning}
share the transitive relevance of \process states. However, we further
constrain the partitioning in order to guarantee the existence of
exactly one consistent partitioning that \processes eventually converge
to.

\begin{definition}[\underline{C}onsistent \underline{p}artitioning (CP)]
  We associate each edge $\langle x, y \rangle \in E$ with a pair of
  positive \underline{w}eights $w_{xy}$ and $w_{yx}$. The shortest
  \underline{p}ath $\Pi_{xz}$ has a distance as a sum of weights
  $\sum\{w_{pq} | \langle p, q \rangle \in \Pi_{xz}\} = |\Pi_{xz}|$
  that is smaller than any other path, with $\Pi_{xx}$ being the
  greatest lower bound of $x$. The only consistent partitioning is a
  set of logical partitions $P_{s\in S}$ such that every \process
  belongs to the partition of its closest source:
  $\forall x \in P_{s}: \nexists P_{s'}$ such that
  $|\Pi_{s'x}| < |\Pi_{sx}|$.
\end{definition}

Unfortunately, \processes do not share a common global knowledge of
the network state. For \processes to reach consistent partitioning
together, a \Process $s$ \underline{a}dding a partition must send a
notification $\alpha_s$ to every \process that is closer to it than
any other source. Since epidemic dissemination and scoped broadcast
have well-aligned objectives, we assume implementations relying on
message forwarding from neighbor-to-neighbor.

\begin{definition}[\label{def:forwarding}Forwarding]
  A \process $x$ \underline{f}orwards a message $m$ by
  \underline{s}ending it to its neighbors:
  $f_x(m) \implies \forall \langle x, y\rangle \in E:
  \TODO{s_{xy}(m)}$.
\end{definition}

\begin{theorem}[\label{theo:fb}\underline{F}orwarding of \underline{B}est (FB)
    $\implies\eventually$CP]
    % 
    Assuming reliable communication links ($s_{xy}(m) \iff \eventually
    r_{yx}(m)$), and a total order on messages ($<$), \processes
    eventually reach consistent partitioning if each \process delivers
    its best message among \underline{r}eceived messages $R_x$, and
    forwards its best message among \underline{d}elivered messages
    $D_x$.
    % , \ie
    % the message $\alpha_s^d$ with the smallest distance $d$ to the
    % source $s$
    % ($(\alpha_s^d = \min R_x) \implies (d_x(\alpha_s^d) \wedge
    % f_x(\TODO{\alpha_s^d}))$).
\end{theorem}

>>>>

This definition encompasses more specific definitions of related
work~\cite{hsiao2005scoped, lue2006scoped, wang2015prodiluvian}. It
underlines the transitive relevance of messages. It also highlights
that the functioning of epidemic propagation is well-aligned with the
objectives of scoped broadcast. As consequence, we assume
implementations relying on the forwarding of messages from
neighbor-to-neighbor.

\begin{definition}[\label{def:forwarding}Forwarding]
  A \process $x$ \underline{f}orwards a message $m$ by
  \underline{s}ending it to its neighbors with metadata that depends
  on its local knowledge: $f_x(m) \implies \forall \langle x, y\rangle
  \in E: s_{xy}(m \oplus \sigma_x)$.
\end{definition}

We use scoped broadcast to efficiently modify the state of each
\process depending on the partitions that exist in the system.

@@

At any time, a \process can decide to become a \emph{source}, hence
creating a new partition in the system by executing an \texttt{Add}
operation. This partition includes at least its source plus
neighboring \processes that are closer to this source than any other
one. Such a distance (or \emph{weight}) is application-dependant: in
the context of maintaining distributed indexes, this would be about
link latency that \nodes could monitor by aggregating \texttt{ping}s;
or operational costs when dealing with multiple tenants.

@@

%% \begin{definition}[\label{def:partitioning}Partitioning]
%%   Let $S \subseteq V$ be the set of \underline{s}ources, and $P_{s\in
%%     S}$ be the \underline{p}artition including at least \Process~$s$,
%%   each \process belongs to at most one partition $\forall x, y \in V,
%%   \forall s,s' \in S: (x \in P_{s} \wedge y \in P_{s'}) \implies (x \neq
%%   y \vee s = s')$, and there exists at least one path $\pi_{xs}$ of
%%   \processes that belong to this partition $\forall z \in \pi_{xs}: z
%%   \in P_s$.
%% \end{definition}

%% Definition~\ref{def:scoped} and Definition~\ref{def:partitioning}
%% share the transitive relevance of \process states. However, we further
%% constrain the partitioning in order to guarantee the existence of
%% exactly one consistent partitioning that \processes eventually
%% converge to.

\begin{definition}[\underline{C}onsistent \underline{p}artitioning (CP)]
  Assuming a set of \underline{s}ources $S\subseteq V$, a positive
  \underline{w}eight $w_{xy}$ associated with each edge $\langle x, y
  \rangle \in E$, we define consistent partitioning as a set of
  logical partitions $P_{s\in S}$ where each \node $x$ belongs to the
  partition of its closest source $s$, \ie there exists a
  \underline{p}ath $\pi_{sx}$ with a sum of weights $|\pi_{sx}| =
  \sum\{w_{pq} | \langle p, q \rangle \in \pi_{sx}\}$ smaller than any
  other path, with $|\pi_{xx}|$ being $x$'s greatest lower bound.
  %$\forall x \in
  % P_{s}:  \nexists P_{s'}$ such that $|\pi_{s'x}| < |\pi_{sx}|$.
\end{definition}

Unfortunately, \processes do not share a common global knowledge of
the network state. For \processes to eventually ($\eventually$) reach
consistent partitioning, each source $s$ must send notifications
$\alpha_s$ to all \processes that are closer to it than any other
source, along with enough control information to allow them to decide
of their closest source.

\begin{theorem}[\label{theo:fb}\underline{F}orwarding of \underline{B}est (FB)
    $\implies\eventually \textup{CP}$]
    % 
    Assuming reliable communication links ($s_{xy}(m) \iff \eventually
    r_{yx}(m)$), and a total order on messages ($<$), \processes
    eventually reach consistent partitioning if each \process delivers
    its best message among \underline{r}eceived messages $R_x$, and
    forwards its best message among \underline{d}elivered messages
    $D_x$.
    % , \ie
    % the message $\alpha_s^d$ with the smallest distance $d$ to the
    % source $s$
    % ($(\alpha_s^d = \min R_x) \implies (d_x(\alpha_s^d) \wedge
    % f_x(\TODO{\alpha_s^d}))$).
\end{theorem}
